\documentclass[x11names,reqno,14pt]{extarticle}
\input{preamble}
\usepackage[document]{ragged2e}
\usepackage{epsfig}

\pagestyle{fancy}{
	\fancyhead[L]{Fall 2022}
	\fancyhead[C]{220A - Groups}
	\fancyhead[R]{John White}
  
  \fancyfoot[R]{\footnotesize Page \thepage \ of \pageref{LastPage}}
	\fancyfoot[C]{}
	}
\fancypagestyle{firststyle}{
     \fancyhead[L]{}
     \fancyhead[R]{}
     \fancyhead[C]{}
     \renewcommand{\headrulewidth}{0pt}
	\fancyfoot[R]{\footnotesize Page \thepage \ of \pageref{LastPage}}
}
\newcommand{\pmat}[4]{\begin{pmatrix} #1 & #2 \\ #3 & #4 \end{pmatrix}}
\DeclareMathOperator{\Perm}{Perm}

\title{220A - Groups}
\author{John White}
\date{Fall 2022}


\begin{document}

\section*{Lecture 1}

Algebra consists, generally speaking, of Groups, Rings, and Fields. These are covered, respectively, in 220A, 220B, and 220C. 


We will begin with the minimal possible axioms, and go on to more extensive axioms. If you start with very few axioms, there are too many examples. If you go too far with axioms, you have too many examples. Group theory is all about binary operations. 

\defn 

Let $S$ be a set. A \underline{binary operation} on $S$ is a function $f:S\times S \to S$. The image of $(a, b) \in S \times S$ under $f$ is often denoted by $a * b$, or simply $ab$. 

\defn
 
A binary operation is called \underline{associative} if, for each $a, b, c \in S$, we have 
\[
(ab)c = a(bc)
\]

\defn

An element $e \in S$ is an \underline{identity} if $ex = xe = x$ for all $x \in S$

\defn 
A \underline{monoid} is a set $S$ equipped with a binary operation which is associative, and which admits an identity. 

\section*{Lecture 2}

Given $x_1, \dots, x_n \in M$, with $M$ a monoid, we may inductively define
\[
\prod_{i=1}^nx_i \eqdef(\prod_{i=1}^{n - 1}x_i)x_n
\]

Fact: In a commutative monoid, given any bijection $\psi:\{1,\dots,n\}\to\{1,\dots,n\}$, 
\[
\prod_{i=1}^nx_{\psi(i)} = \prod_{i=1}^nx_i
\]

\defn

A \underline{group} $G$ is a monoid such that $\forall x \in G$, $\exists y \in G$ such that $xy = yx = e$. Here, $y$ is called an inverse of $x$. Note, we have not yet proved that $y$ is unique. 

\claim

For any $x \in G$, for $G$ a group, $x^{-1}$ is unique. 

\proof

Let $y, y'$ be inverses to $x$. Then 
\[
y = ey = (y'x)y = y'(xy) = y'e = y'
\]

\qed

\exm

Let $G$ be a group. Let $S$ be a nonempty set. 

The set of maps from $S$ to $G$, denoted $M(S, G)$, is a group as follows. 

Given $f, g \in M(S, G)$, define $fg(x) = f(x)g(x)$. 

Identity: $e(x) = e \forall x \in S$. 

Inverses: For an $f \in M(S, G)$, $f^{-1}$ is the function such that $(f^{-1})(x) = f(x)^{-1}$. 

\exm

Given a nonempty set $S$, define $\operatorname{Perm}_S$ as the set of all bijective maps from $S$ to $S$. This forms a group under composition. 

When $|S|$ is a finite set of size $n$, we identify $\operatorname{Perm}_S$ with $S_n$. 

\exm

Let $k$ be a field, and $V$ a vector space over $k$. Then $GL(V)$ and $GL(n, k)$ are groups. 

\exm

$\Q$ is a group. $\Q^\times$, which is $\Q \backslash\{0\}$, is a group under multiplication. 

This works for any field. 

\exm

$\Z$ is an additive group. $\{1, -1\}$ forms a multiplicative group. 

\defn

A group $G$ is called \underline{cyclic} if $\exists a \in G$ such that every $x \in G$ is of the form $a^n$ for some $n \in \Z$. Such an $a$ is called a generator. 

So $\Z$ is (additively) cyclic with generators $\pm 1$.

\defn

The \underline{order} of $G$ is $|G|$ if $G$ is finite, and $\infty$ otherwise. 

\defn

Let $N \in \N$. Then the $N$th \underline{roots of unity} form a multiplicative subgroup of $\C$. It is cyclic, and generated by $e^{\frac{2\pi i k}{N}}$. 

An element $e^{\frac{2\pi ik}{N}}$ is called a primitive if it is a generator. 

For example, for $N = 4$, our group would be $\{\pm 1, \pm i\}$. The primitives would be $\pm i$. 

$\{e^{\frac{2\pi i k}{N}}\}$ is a generator of the $N$th roots of unity if and only if $k$ is coprime to $N$. 

\subsection*{}

Let $P$ be a regular polygon with $N$ sides. Let $\Aut(P) = \{$ automorphisms of $P\}$. The automorphisms of $P$ form a group, and it is generated by a reflection and a rotation. 

\defn

A subset $S \subseteq G$ generates $G$ if every $x \in G$ can be written $x = s_1^{a_1}\cdots s_n^{a_n}$ with $s_i \in S$ for each $i$. 

Let $\sigma, \tau \in \Aut(P)$. We can see that $\tau\sigma\tau^{-1} = \tau\sigma\tau = \sigma^{-1} = \sigma^{n - 1}$. 

If $G$ is generated by $S$, any expression of the form $s_1^{a_1}\cdots s_n^{a_n}$, with $s_i \in S$ and $a_i \in \Z$ is called a word in the generators. 

If $W$ is a set of words in $S$, then 
\[
\langle S \mid \rangle
\]
is the free group on $S$ quotiented by the normal subgroup generated by all elements of $W$. This is called a group presentation of the resulting group. 

So $\Aut(P) = \langle \sigma, \tau, \sigma^n = \tau^2 = (\tau\sigma)^n = e\rangle$

\section*{Lecture 3}

\defn

A monoid $M$ is called \underline{commutative/abelian} if  for all $x, y \in M$, $xy = yx$. 

$N\subset M$ is called a \underline{submonoid} iff $x,y \in N\implies xy \in N$, and $e \in N$. 

A function $f:M\to N$ is a \underline{homomorphism of monoids} if $f(xy) = f(x)f(y)$ for all $x, y \in M$. 

\defn

A \underline{commutative/abelian group} and a \underline{homomorphism of groups} are defined exactly as above. A \underline{subgroup} is defined similarly as above, but with the added requirement that $x\in N$ implies $x^{-1}\in N$. 

\exm

Let $\mu_N$ denote the $N$th roots of unity, generated by $e^{\frac{2\pi i}{N}}.$ We have a morphism $f:\Z\to\mu_N$ given by $f(k) = e^{\frac{2\pi ik}{N}}$. 

\exm

We have a $\pi:D_n\to\mu_N$ given by $\pi(\sigma)= -\pi(\tau) = -1$, $\pi(e) = e$. 

\defn

Given $\phi:G\to H$ a homomorphism of groups, the \underline{kernel} of $\phi$, denoted $\ker(\phi)$, is defined by
\[
\ker(\phi) \eqdef \{g \in G \mid \phi(g) = e \}
\]
Similarly, we define the \underline{image}, $\Im(\phi)$, by 
\[
\Im(\phi) = \{h \in H \mid \exists g \in G, \phi(g) = h \}
\]

\defn

If $N$ is a subgroup of $G$, then we say that $N$ is \underline{normal} if for all $g \in G$ and $n \in N$, we have $gng^{-1} \in N$. 

\lem Given a morphism $\pi: G\to H$, then 
\begin{enumerate}
\item $\pi$ is injective if and only if $\ker(\pi) = \{e\}$. In this case, we say $\pi:G\hookrightarrow H$. 
\item $\pi$ is surjective if and only if $\Im(a) = H$. In this case, we say $\pi:G\Rightarrow H$. 
\end{enumerate}

\proof

The second assertion is obvious, but we shall prove the first. Suppose that $\pi$ is injective. Then clearly $\ker(\pi) = \{e\}$. Now, suppose that $\ker(\pi)$ is trivial. Suppose that $\pi(a) = \pi(b)$. Then $\pi(a)\pi(b)^{-1} = e = \pi(ab^{-1})$, so $ab^{-1} = e$, so $a = b$. 

\qed

\prop

Given a group homomorphism $\pi: G\to H$, the kernel of $\pi$ is a normal subgroup of $G$. 

\proof

Choose $x \in G, y \in \ker(\pi)$. Then 
\[
\pi(xyx^{-1}) = \pi(x)e\pi(x)^{-1} = e
\]
So $xyx^{-1} \in \ker(\pi)$. 

\defn

Given $H\unlhd G$, we define the \underline{quotient group} as follows. 

For any $g \in G$, we denote the left coset of $H$ containing $g$ as $gH$. The set of left cosets forms a group, with the group operation given by 
\[
(gh)(g'H) = (gg')H
\]

This is not, a priori, well defined, so we will prove that now. That is, 

\lem 

$gH = g'H$ if and only if $g(g'^{-1}) \in H$. 

\proof

First, assume $gH = g'H$. Then $g \in g'H$, i.e there is some $h \in H$ such that $g = g'h$. Thus, $gg'^{-1} = h \in H$. Now, assume $g(g'^{-1}) \in H$. Then choose $x \in gH$. There is an $h \in H$ such that $g = g'h$. 

Then $g'g^{-1}x = g'g^{-1}gh = g'h$, and so $x = g'h\underbrace{(g'g^{-1})}_{\in H}\in g'H$.

This is one inclusion, and the other follows similarly. 

\claim If $H$ is normal, then $G/H$, the set of left cosets of $H$, inherits a group structure from $G$ as follows: $(gH)(g'H) \eqdef gg'H$

\proof

For well-definedness, assume that $gH = ghH$. Note that $(ghg')(gg')^{-1} = ghg'^{-1} \in H$. So by a previous claim, $gH = ghH$.



\section*{Lecture 4}


\defn

Let $G$ be a group. The \underline{automorphism group} $\Aut(G)$ of $G$ is defined as the group of bijective homomorphisms $f:G\to G$, with the group operation defined as function compositions. 

For a given $x \in G$, there is an automorphism given by $g \mapsto x^{-1}gx$, which is conjugation by $x$. Any automorphism which can be given by conjugation is called an \underline{inner automorphism}, and they form a normal subgroup of $\Aut(G)$. 

\claim Consider $f:G\to\Aut(G)$ defined by $x \mapsto (g \mapsto x^{-1}gx)$. This is a group homomorphism. 

\proof

For any $g_1, g_2 \in G$, and any $x \in G$, we have

\[
f_x(g_1g_2) = x^{-1}g_1g_2x = x^{-1}g_1xx^{-1}g_2x =f_x(g_1)f_x(g_2)
\]

So, $f_x \in \Aut(G)$. Now, let $x, y \in G$ and $g \in G$. Then

\[
f_{xy}(g) = (xy)^{-1}gxy = y^{-1}x^{-1}gxy = y^{-1}(x^{-1}gx)y = (f_x \circ f_y) (g)
\]

\qed

\defn 

The \underline{opposite group} of a group $G$ has the same underlying set as $G$, whose group operation (denoted $*_{op}$) is defined by

\[
(x*_{op}y) \eqdef yx
\]

\claim This is associative

\proof

We have 

\[
(x*_{op}y)*_{op}z = (yx)*_{op}z = zyx = x*_{op}(zy) = x*_{op}(y*_{op}z)
\]

\qed

Fact: $H$ normal is equivalent to every left coset of $H$ is a right coset.

Notation: $H\unlhd G$ means $H$ is a \underline{subgroup} of $G$. 

$H\lhd G$ means $H$ is a \underline{normal subgroup} of $G$. 

\lem Let $K\unlhd H \unlhd Q$. Let $\{x_i\}$ be a set of left coset representatives for $H\unlhd G$, i.e. $\{x_iH\}$ is a complete, non-repeating set of left cosets. 

Let $\{y_i\}$ be a set of left coset representatives for $K\unlhd H$. 

Then $\{y_ix_i\}$ is a set of left coset representatives for $K \unlhd G$. 

\cor

If $(G:H)$ denotes the number of left cosets of $H$ in $G$, then 
\[
(G:K) = (G:H)(H:K)
\]

\proof

We will prove the lemma, and the corollary will follow. 

First, why is it a complete list of cosets? First, if $g \in G$, then $\exists ! x_i$ such that $g \in x_iH$, i.e. $g = x_ih$ for some $h \in H$. Further, for $h \in H$, $\exists!y_j$ such that $h \in y_jK$. Then $g = x_ih \in x_iy_jK$.

Why is it unique? Exercise 

\qed

\defn

$|G| = (G:\{e\})$. 

\cor

For normal subgroups, $|G| = |G/N|\cdot|N|$

\defn

If $g \in G$, then the \underline{order of the element $g$} is defined as the smallest integer $k$ such that $g^k = e$. If there is no such $k$, then we say the order is $\oo$. 

\section*{Lecture 5}

Recall: $(G:H)$ is the index of $H$ in $G$, which is the number of left cosets of $H$ in $G$. We know that
\[
\frac{|G|}{|H|} = \frac{(G:e)}{(H:e)}
\]

\cor

If $g \in G$, $o(g) < \oo$, $|G|<\oo$, then $o(g) \mid |G|$. 

\cor If $G$ is a group with $|G| = p$, a prime number, then $o(g) = 1$ or $p$ for all $g \in G$. 

\lem

There are exactly 1 element of order 1 in any group. 

\proof

We know $o(e) = 1$. If $e, e'$ are both units in $G$, then $e =ee' = e'$. 

\cor

If $|G| = p$, then $G$ is cyclic. 

\proof Choose $g \neq e$. Then $o(g) = p$, so $\{g^0,g^1,\dots, g^{p - 1}\}$ is $p$ distinct elements in $G$, so $g$ must generate $G$. 

Classification of groups of order $n$. 

\begin{align*}
n = 1 & &G = \{e\} \\
n = 2 & &\text{ prime} \\
n = 3 & &\text{ prime} \\
? \\
\end{align*}

We will classify all groups of order $4$ by force, without the proper tools (eg semidirect products, sylow theorems). 

If there exists a $g$ such that $o(g) = 4$, then $G$ is cyclic. Otherwise, all $g \neq e$ have order 2. Let $G = \{e, a, b, c\}$. If $ab = a$, then $b = c$, and similarly if $ab = b$. So $ab = c$. 

Thus, we know every group of order $4$. 

\begin{center}
\begin{tabular}{c | c c c c }
 & e & a & b & c \\
\hline
e & e & a & b & c \\
a & a & e & c & b \\
b & b & c & e & a \\
c & c & b & a & e \\
\end{tabular}
\end{center}
The above is the \it{Cayley table} of the group. 

\claim The $G$ above, the \underline{Klein $4$-group}, is isomorphic to $\Z/2\Z \times \Z/2\Z$. 

\defn

For $H, K$ groups, we define their direct product $H\times K$ by 
\[
H\times K = \{(h, k) \mid (h,k)(h',k') = (hh', kk')\}
\] 

Continuing with the classification, 5 is prime, so we move on to $n = 6$. If it's non-cyclic, then $o(g) \in \{1, 2, 3\}$ for all $g \in G$. 

We can classify them by setting up a semidirect product...

Let $H, K$ be groups, and let $\psi:H\to\Aut(N)$. We define the semidirecN product by $H\ltimes N$. The underlying set is $H\times N$, and the group operation is given by 
\[
(x_1,h_1)(x_2,h_2) = (x_1\psi(h_2)(x_2),h_1h_2)
\]




$\vdots$ 

7 is prime. Consider $n = 8$. 

The abelian groups of this order are $\Z/8\Z$, $\Z/4\Z\times\Z/2\Z$, and $\Z/2\Z\times\Z/2\Z\times\Z/2\Z$.  

If it isn't abelian, it could be $D_4$ or $Q_8$. $Q_8$ is the quaternion group, consisting of $\{\pm 1, \pm i, \pm j, \pm k\}$, with $i^2 = j^2 = k^2 = ijk = -1$. 

It is called this because it is the group of units of $\H$, the quaternions. $H$ is for William Rowe Hamilton, who first wrote this down, and whose name scans exactly the same as ``Alexander Hamilton." 

Let's move on to $n = 9$. It could be cyclic, and if it is cyclic but non-ableian, then it could be $\Z/3\Z\times\Z/3\Z$. It could be that there exists a different group such that $g\neq e \implies o(g) = 3$. This turns out not to be the case, but proving it with only what we have now would be tough. 

If $|G| = 10$, then $G$ is either cyclic, a product of cyclics, or dihedral. 

\underline{Basic Constructions}
\begin{enumerate}
\item Theorem about finitely generated abelian groups
\item Semidirect product constructions: If $N\lhd G$, then $G$ acts on $N$ by conjugation: $\psi:G\to\Aut(N)$, $x\mapsto(g\mapsto xgx^{-1})$. 

For a semidirect product, we need $H \lhd G$ such that $|G| = |N||H|$. 
\item Special constructions.  
\end{enumerate}

One of the big steps for the class project will be finding normal subgroups. 

\defn A group $G$ is \underline{simple} if it has no nontrivial normal subgroups. 

Strategy to understand all finite groups: 
\begin{enumerate}
\item Find simple ones. This one has been achieved through work starting in the 1960s through the 1980s, and 10000 journal pages. 
\item Assemble them into more complicated ones
\end{enumerate}

There are a few big classes of finite simple groups: 
\begin{itemize}
\item Alternating group $A_n$, for $n \geq 3$
\item Finite groups of Lie type, groups of matrices over finite fields.  
\end{itemize}

There are also the sporadic groups, of which there are 26, which don't fit into any other category. They include the monster group, the largest simple group, and the baby monster group, the second largest simple group. 

\section*{Lecture 5}

Correction to a statement made last time: 

There is something called the $J$-invariant of elliptic curves. An elliptic curve is a torus with a complex structure. This is $\C/\Lambda$, where $\Lambda$ is a free $\Z$-module of rank $2$. Let $\tau$ be one of the generators. We want all things of the form $\Lambda_\tau = \{m + n\tau\}$. 

For a change of basis in $\Lambda$, $\begin{pmatrix} a & b \\ c & d \end{pmatrix} \in \SL_2(\Z)$. I don't really know what the heck he's talking about lol, I'm gonna go fiddle with the preamble. 

Groups can be represented on vector spaces. We want to define ``multiplication by elements of $G$" on elements of $V$, such that $g\cdot(v_1 + v_2) = g\cdot v_1 + g\cdot v_2$, nad $(g\cdot(cv)) = c(g\cdot v)$, where $c \in k$, the ground field of $V$. 

This gives a homomoprhism $G\to \GL(V)$. 

More generally, groups can ``act" on sets (ignoring any other structure). Recall that for a set $S$, we define $\operatorname{Perm}(S)$ as the set of all bijective $f:S\to S$. 

\defn

A \underline{group action} on the set $S$ is a group homomorphism $\pi:G\to\operatorname{Perm}(S)$. We denote for $g \in G$, $s \in S$, we denote $\pi(g)(s)$ by $g\cdot s$. Note that $(g_1\cdot g_2)\cdot s = g_1 \cdot (g_2 \cdot s)$. 

\defn

If $G\to\operatorname{Perm}(S)$ is a group action, the \underline{orbit} of the action containing $s \in S$ is 
\[
\{g\cdot s \mid g \in G\}
\]
Note that $S = S_1\coprod S_2 \coprod \cdots$ is a decomposition of $S$ into disjoint orbits. We can then obtain $G\to \operatorname{Perm}(S_i)$. 

Consider $n\times n$ matrices that are invertible and have a single $1$ and all the rest $0$ in each row and each column. This is a subset of $\GL(n, \mathbb{F})$. We can represent a group as matrices by constructing a basis indexed by the group elements $\{v_{g_1}, v_{g_2}, \dots, v_{g_d}\}$, and define $\pi(g)(v_{g_i}) = v_{gg_i}$. 

In this way we can identify $S_3$ with the set of $3\times 3$ matrices which act on the basis $\{v_{s_1}, v_{s_2}, v_{s_3}\}$ according to the permutation of the indices. 

So, given $\sigma \in jS_3$, there is a $\pi(\sigma)\in \GL(3, \mathbb{F})$, and $(6, \operatorname{char}\mathbb{F})$. 

Any element of $S_n$ can be decomposed into cycles, like so
\begin{align*}
S_n = & \{1, \sigma(1), \sigma^2(1), \dots, \sigma^n(1) \} \\
	\cup	& \{i, \sigma(i), \dots \} \\
	\cup	& \vdots \\
\end{align*}

\section*{Lecture 6}

Recall: $S_n \eqdef \Perm\{1, 2, \dots, n\}$. 

For $\sigma\in S_n$, $\langle\sigma\rangle$ denotes the cyclic subgroup of $S_n$ generated by $\sigma$.

We arrange elements such that $\sigma(i_k^{(j)}) =\begin{cases} i_{k + 1}^{(j)} & k < n_j \\ i_1^{(j)} & k = n_j \\ \end{cases}$

\[
\{1,2,\dots,n\} = \coprod_j\{i_1^{(j)}, \dots, i_{n_j}^{(j)}\}
\]



So $\sigma = [i_1^{(1)}, \dots, i_{n_1}^{(1)}]\cdots[i_1^{k)},\dots,i_{n_k}^{(k)}]$

Let $G \subset S_4$, with $G = \{e, [12][34], [13][24], [14][25]\}$. We can see $G \cong V_4$, the Klein $4$-group. 

\claim

$G$ is normal in $S_4$. 

\proof

Let $\sigma = [i_1^{(1)}, \dots, i_{n_1}^{(j)}]\cdots$. Then
\[
\tau^{-1}\sigma\tau = [\tau(i_1^{(1)}), \dots, \tau(i_{n_1}^{(j)})]\cdots
\]

for $\tau \in S_4$. Indeed, $\tau^{-1}([12][34])\tau = [\tau(1)\tau(2)][\tau(3)\tau(4)] \in G$

\qed

Let $S_n$ act on the $K$, the functions from $\Z^n$ to $\Z$. So this is $\pi:S_n\to\Perm\Hom_{Set}(\Z^n, \Z)$. 

Define $\pi(\sigma)f(x_1, \dots, x_n) = f(x_{\sigma(1)}, \dots, x_{\sigma(n)})$. It is clear that $\pi(\tau)\pi(\sigma)f = \pi(\tau\sigma)f$. 

$\coprod_j\{1_i^{(j)},\dots,i_{n_j}^{(j)}\}$, 

Here is Quinn's explanation of how this cycle/orbit nonsense notation works. 

\[
\tau(\sigma(\tau^{-1}(x)))
\]

\[
\tau^{-1}(x) = i_k^{(j)} \implies x = \tau(i_k^{(j)})
\]

\[
\sigma(\tau^{-1}(x)) = i_{k +1}^{(j)}
\]

\[
\tau(\sigma(\tau^{-1}(x))) = \tau(i_{k + 1}^{(j)})
\]

Let $f(x_1, \dots, x_n) = \prod_{1\leq i<j\leq n}(x_i - x_j)$. 

Define $\varepsilon:S_n\to\{\pm 1\}$ by $\pi(\sigma)f(x_1, \dots, x_n) = \varepsilon(\sigma)f(x_1, \dots, x_n)$. 

\prop

First, every $\sigma$ is a product of simple transpositions. Second, $\sgn(\varepsilon) \eqdef (-1)^{\text{number of transpositions}}$

\proof
\qed

$\ker\sgn = A_n$, the alternating group. 

We see $G\lhd A_4$, with $|A_4/G| = 3$. So $A_4/G \cong Z_3$. 

\thm For later: $A_n$ is simple for $n\geq5$. 

\defn

A \underline{tower of subgroups} is a collection of subgroups 

\[
G = G_0\supseteq G_1\supseteq G_2 \supseteq \cdots \supseteq G_m = \{e\}
\]

A tower is \underline{normal} if $G_{i + 1}\lhd G_i$. 

A normal tower is \underline{ableian} (\underline{cyclic}) if $G_i/G_{i + 1}$ is abelian (cyclic) for all $i$.

A group $G$ is \underline{solvable} if it has an abelian tower with $G_m = \{e\}$. 

A \underline{refinement} of a tower is a new tower into which some groups have been inserted. 


\prop

Any abelian tower has a cyclic refinement. 

\section*{Lecture 7}

\lem

Let $G$ be a finite abelian group. Then $G$ admits a cyclic tower ending in $\{e\}$. 

\proof

We will induct on $|G|$. If $G = \{e\}$ it's true. This is our base case. 

Otherwise, let $x \in G$ be a non-identity element of $G$. Let $X = <x> \lhd G$.

Consider $G' = G/X$. $|G/X|<|G|$, so 
\[
G/X = G'_0 \supseteq G_1' \supseteq \cdots \supseteq \{e\}
\]

Let $G_j = \pi_X^{-1}(G_j')\lhd G$, where $\pi_X:G\to G/X$ is the projection map. 

Then

\[
G \supseteq G_0 \supseteq \cdots \supseteq X
\]
This is a cyclic tower, because $G_j/G_{j + 1} = G'/G_{j + 1}'$. We can add $\{e\}$ to the end of this tower, because $X$ is cyclic. 

\prop

Any abelian tower admits a cyclic refinedment. 

\proof

Consider

\[
\Gamma = \Gamma_0 \supseteq \Gamma_1\supseteq \cdots\supseteq \Gamma_i\supseteq \cdots
\]

We know $\Gamma_{i + 1}\lhd \Gamma_i$ such that $\Gamma_i/\Gamma_{i + 1}$ is abelian. 

$\Gamma_i/\Gamma_{i + 1}$ has a cyclic tower by the lemma; so the inverse image of $\Gamma_{i + 1}$ in $\Gamma_i$ under the projection map has a cyclic tower, so 
\[
\Gamma_i\underbrace{\supseteq\cdots\supseteq}_{\text{new cyclic tower}}\Gamma_{i + 1}
\]

\exm

Let $G = \GL(n, \mathbb{F})$. 

Let $N$ be the strictly upper triangular matrices, i.e. upper triangular matrices with 0 on the diagonal. 

\[
\begin{pmatrix} 0 & a_1 & a_2 & \cdots & a_{1n} \\
					 0 & 0 & \cdots & \cdots & a_{2n} \\
					\vdots & \ddots & \ddots & \ddots & \vdots \\
					0 & \cdots & \cdots & 0 & a_{nn} \\
\end{pmatrix}
\]

$N^{r - 1}$ has the form 

\[
\begin{pmatrix} 0 & 0 & a_{12} & \cdots & a_{1n} \\
					 0 & 0 & \cdots & \cdots & a_{2n} \\
					\vdots & \ddots & \ddots & \ddots & \vdots \\
					0 & \cdots & \cdots & \cdots &0 \\
\end{pmatrix}
\]

Let $U_r = \{I + A \mid A \in N_r\}$. This is a subgroup of $\GL(n, \mathbb{F})$. 

We have a tower
\[
U_1\supseteq U_2 \supseteq \cdots \supseteq U_k
\]

It turns out that $U_{i + 1}$ is normal in $U_i$, but this is not obvious. We have that $U_i/U_{i + 1} = \mathbb{F}^{n - i}$. 

Let's review some things we know. $S_1$ is trivial, $S_2$ is cyclic, $S_3\unlhd A_4 \cong \Z/3\Z$. $S_3 /A_3 \cong \Z/2\Z$, so $S_3$ is solvable. 

$S_4 \unlhd A_4 \unlhd \Z/2\Z \times \Z/2\Z$. We know $A_4 /(\Z/2\Z\times\Z/2\Z) \cong \Z/3\Z$, and $S_4/A_4 \cong \Z/2\Z$, so $S_4$ is solvable. 

For all $n$, $S_n\unlhd A_n$ and $S_n/A_n \cong \Z/2\Z$. But for $n \geq 5$, $A_n$ is simple, and so $S_n$ is not solvable. 

\subsection*{\underline{Group actions}}

Ways of making a group act on itself:
\begin{enumerate}
\item Conjugation: For $x \in G$, there is a map $c_x:G\to G$ defined by $c_x(y) = xyx^{-1}$. 
\item $G$ acts on subgroups which are conjugate to the subgroup $H$: $c_x(H) = xHx^{-1}$. 
\item Left translation: Given $x \in G$, we can define $T_x\in \Perm(G)$. This is defined by $T_x(y) = xy$. We can define right-translation similarly. 
\end{enumerate}

If $G$ acts on $S$, meaning we have a homomorphism $G\to\Perm(S)$. This is sometimes written ${}^G\circlearrowright S$. 

\defn

Suppose ${}^G\circlearrowright S$. Then, for $s \in S$, $Gs = \{gs \mid g \in G\}$ is the \underline{orbit} containing $s$. 

Similarly, $G_s = \{g \in G \mid gs = s\}$ is the \underline{stabilizer} of $s$. 

\claim 

$G/Gs$ and $Gs$ are in 1-1 correspondence. 

\proof

If $x, y \in G$, then $xG_s = yG_s$ is equivalent to $y^{-1}x G_s = G_s$, which is equivalent to $y^{-1}x \in G_s$, which is equivalent to $yxs = s$, which is equivalent to $xs = ys$, which is equivalent to $x, y$ in the same orbit. 

\exm

$S_3$ acts on $S = \{1, 2, 3\}$. 
\begin{align*}
G_1 & = \{e, (23)\} \\
G/G_1 & = \{eG_1, (12)G_1, (123)G_1, (132)G_1\} \\
\end{align*}

\section*{Lecture 8}

\thm $A_n$ is simple for $n \geq 5$. 

\lem

\begin{enumerate}[label=(\alph*)]
\item $A_n$ is generated by $3$-cycles
\item If $n \geq 5$, all $3$-cycles are conjugate in $A_n$ 
\end{enumerate}

\proof

\begin{enumerate}[label=(\alph*)]
\item We want to consider the product $[ij][rs]$. If $i, j, r, s$ are not all distinct, then either 

\begin{itemize}
\item $[ij][rs] = \Id$ if $i = r$ or $j = s$, or vice versa
\item $[ij][rs]$ is a $3$-cycle
\item $[ij][rs] = [isj]$
\end{itemize}

Or, $[ij][rs] = [ijr][rs]$ if they are all distinct. 


\item Let $\gamma \in S_n$. We have
\[
\gamma[i_1\cdots i_m]\gamma^{-1} = [\gamma(i_1)\cdots\gamma(i_m)]
\]
Given $[ijk]$ and $[i'j'k']$, let $\gamma(i) = i'$, $\gamma(j) = j'$, and $\gamma(k) = k'$. Then $\gamma[ijk]\gamma^{-1} = [i'j'k']$. If $\gamma$ is even we are done. If $\gamma$ is odd, then there exist distinct $r, s$ distinct from all of $i, j, k$ $(n \geq 5)$, then $\gamma' = \gamma[rs]$.
\end{enumerate}

\proof of theorem. 

Let $N \neq \{e\}$ be a normal subgroup of $A_n$. Let $\sigma \in N$, $\sigma\neq e$, have the maximum number of fixed points among all such $\sigma$s. So some orbit of $\sigma$ has $> 1$ element.

Suppose all orbits of $\sigma$ have $1$ or $2$ elements. Then there must be at least $2$ orbits with two elements each. Otherwise, $\sigma = [ij]\not\in A_n$

So $\{1, \dots, n\} = \mc{O}_1\coprod \mc{O}_2 \coprod \cdots \coprod \mc{O}_k$.  

On two such orbits, $\sigma$ acts as $\sigma = [ij][rs]$ on $\{i,j,r,s\}$. Let $k \neq i,j,r,s$ (again we use the assumption $n \geq 5$). Let $\tau = [rsk]$, and $\sigma' = \tau\sigma\tau^{-1}\sigma^{-1} = [\tau,\sigma] \in A_n$. Now, $\sigma ' = \tau\sigma\tau^{-1}\sigma^{-1}\in N$.

\begin{align*}
\sigma'(i) & = \tau\sigma\tau^{-1}\sigma^{-1}(i) \\
			  & = \tau\sigma\tau^{-1}(j) \\
			  & = \tau\sigma(j) \\
			  & = \tau(i) \\
			  & = i \\
\end{align*}

So

\begin{align*}
\sigma'(j) & = \tau\sigma\tau^{-1}\sigma^{-1}(j) \\
			  & = \tau\sigma\tau^{-1}(i) \\
			  & = \tau\sigma(j) \\
	        & = \tau(j) \\
			  & = j \\
\end{align*}

So if $\sigma(x) = x$, and $x \neq i, j, r, s$, then $\sigma'(x) = x$, so $\sigma'$ has more fixed points than $\sigma$. 

Hence, $\sigma$ is an orbit with at least $3$ elements, $i, j, k, \dots$. 

If $\sigma \neq [ijk]$, then $\sigma$ must move at least $2$ other numbers $r, s$. Let $\tau = [krs]$, and again define $\sigma' = [\tau, \sigma] \in N$. $\sigma'(i) = i$ if $\sigma(x) = x$, but $x \in \{k,r,s\}$. Then $\sigma(x') = x'$. 

So $\sigma'$ again has more fixed points than $\sigma$, contradicting our assumption that $\sigma \neq [ijk]$. 

Hence $[ijk] \in N$. 

So all $3$-cycles are in $N$, because $N \lhd A_n$. Because $A_n$ is generated by 3-cycles, we must conclude $N = A_n$. 
\qed

Enough horrible cycle garbo. 

\subsection*{\underline{Back to group actions}}

\exm

Let $G = \SL_2(\R)$, $S = \ms{H} = \{z \in \C \mid \Im(z) > 0 \}$.

Define 
\[
\begin{pmatrix} a & b \\ c & d \\ \end{pmatrix} \cdot z = \frac{az + b}{cz + d}
\] 
These are the ``fractional linear transformations"

If $\begin{pmatrix} a & b \\ c & d\\ \end{pmatrix} \in \SL_2(\R)$, then 
\begin{align*}
\Im\left(\frac{az + b}{cz + d}\right) & = \Im\left(\frac{(az + b)(c\bar{z} + d)}{|cz + d|^2}\right) \\
					& = \frac{1}{|cz + d|^2}\Im(\cdots )\\
\end{align*}

So 

\[
\frac{(ad - bc)\Im(z)}{|cz + d|^2} = \frac{\Im(z)}{|cz + d|^2}
\]

We want $ai + b = i(ci + a) = -ci + d$. 

So the stabilizer of $i$ is $\{\begin{pmatrix} a & b \\ c & d \end{pmatrix} \in \SL_2(\R) \mid \frac{ai + b}{ci + d} = i \}$. 

Let $K = \{\begin{pmatrix} \cos\theta & \sin\theta \\ -\sin\theta & \cos\theta\end{pmatrix} \}$. $G/K$ is in bijection with $\ms{H}$. 

\subsection*{\underline{Orbits of group actions}}

Suppose ${}^G \circlearrowright S$. Then of course $S = \coprod_{i\in I}G_{s_i}$, where we have chosen exactly one $s_i$ from each orbit. If $S$ is finite, then 
\[
|S| = \sum_{i\in I}|G_{s_i}| = \sum_{i\in I}(G:G_{s_i})
\]

\defn

If $G$ acts on itself by conjugation, $G_x = \{g \in G \mid gxg^{-1} = x\}$, and is referred to as the \underline{normalizer of $x$}. 

Then 
\[
(G:I) = \sum_{x \in C}(G: G_x)
\]
Where $C$ is a set of representations of conjugacy classes. This is often called the ``class equation." 

This can be written

\[
(G:1) = |Z| + \sum_{x\in C, x\not\in Z}(G:G_x)
\]


\section*{Lecture 8}

\section*{Lecture 9}

\subsubsection*{\underline{Direct Product}}

Consider the sequence of sets $\{S_i\}_{i\in I}$. Define
\[
\bigtimes_{i\in I}S_i \eqdef \{(x_i)_{i\in I}\}
\]

The direct product can be defined in general in any abelian category. 

\subsubsection*{\underline{Direct Sum}}

Let $\{G_i\}_{i\in I}$ to be a sequence of abelian groups. Then define

\begin{align*}
\bigoplus_{i\in I}G_i & \eqdef \{(x_i)_{i\in I} \mid x_i \in G, \text{ all but finitely many }x_i = 0\} \\
							  & \subseteq \prod_{i\in I}G_i
\end{align*}

Define $\lambda_j(x) = \begin{cases} x_j & i = j \\ 0 & \text{otherwise} \end{cases}$

\prop

Let $A_j, B$ be abelian groups. Suppose $f_j:A_j\to B$ is a group homomorphism for all $j$. Then there exists a unique $f:\bigoplus A_j \to B$ such that $f \circ \lambda_j = f_j$ for all $j$.

\proof

Define $f((x_i)_{i\in I}) = \sum_{i\in I}f_i(x_i)$, a finite sum. Note such an $f$ being well defined requires the hypothesis that only finitely many $x_i$ are not the identity. 

\defn

A \underline{basis} for an abelian group $A$ is a collection of elements $e_i \in A$ such that every $x \in A$ admits a unique representation in the form 
\[
x = \sum_{i}(x_i \cdot e_i)
\]
where the $x_i$ are integers. If $Z_i \cong \Z$, then $A \cong \oplus_{i \in I}Z_i$. 

We call this the \underline{free abelian group with basis $\{e_i\}$}

More generally, let $S$ be a set. We can define $\Z\langle S\rangle$ as the \underline{free abelian group generated by $S$}, defined as
\[
\Z\langle S \rangle \eqdef \{\phi:S\to\Z \mid \phi(x) = 0\text{ for all but finitely many }x\}
\]

\lem

Let $\begin{tikzcd} A\ar[r, "f"] & A'\end{tikzcd}$ be a surjective homomorphism of finite abelian groups, with $A'$ free, and let $B \subseteq A$ be a subgroup. 

Then there exists a $C \subseteq A$ such that $f|_C:C \to A'$ is an isomorphism, and $A \cong B \oplus C$. 

\proof

Let $\{e_i\}_{i\in I}$ be a basis of $A'$. $f$ is surjective by hypothesis, so we may choose $a_i \in A$ such that $f(a_i) = e_i$ for all $i$. Note $\ker f|_B = B$. The rest of the proof is in lang. lol. 

\thm

Let $A$ be a finitely generated free abelian group with $n$ generated. Let $B \lhd A$ be any subgroup. The $B$ is free with $\leq n$ generators. `

\proof

We will prove this by induction on $n$. 

If $n = 1$, then $A \cong \Z$. We know all of the subgroups of $\Z$: if $B \subseteq A$, then $B = N\Z$ is free, generated by $N$. 

Now for the inductive step: 

$f:A\to \Z\langle x_i \rangle $ be projection. Let $B_1$ be the kernel of $f|_B$. Then $B_1 \subset \Z \langle x_1, \dots, x_{n - 1}\rangle$ is a free group on $\leq n - 1$ generators. Let $C \subset A$ be such that $C \oplus \Z\langle x_i, \dots, x_n\rangle \cong \langle x_1, \dots, x_n\rangle$. Then $B \cong (B \cap C) \oplus B_1$. 

Let $A$ be an abelian group generated by a finite set $S = \{x_1, \dots, x_n\}$. 

Define $\begin{tikzcd} \Z\langle S \rangle \ar[r, "hom"] & A \end{tikzcd}$ by $\phi \mapsto \sum\phi(x_i)x_i$ for $x_i \in A$. 

Let $K = \ker(\Z\langle S \rangle \to A)$ be a subgroup of a free group on $ \leq n$ generators. So it is also free on $\leq n$ generators. 

\defn

Let $f:A\to B$ be a homomorphism of Abelian groups. Then define

\begin{align*}
\ker(f) & = \{a in A \mid f(a) = 0 \} \\
\im(f) & = \{ b \in B \mid b = f(a)\text{ for some }a \in A \} \\
\coker(f) & = B/\im(f) \\
\end{align*}

So we have $\Z[T] \to K$ with cokernel $\Z\langle S \rangle/K \cong A$. 

\subsection*{\underline{Conclusion?}}

Any finitely generated abelian group is the cokernel of a homomorphism $\Z[T]\to \Z[S]$ between finitely generated free abelian groups. 

\defn (Smith normal form for integer matrices)

Let $A$ be a $p \times n$ matrix with entries in $\Z$. We say $A$ is in \underline{Smith Normal Form} if there exist nonzero integers $a_1,\dots, a_n$ such that $a_i \mid a_{i + 1}$, and $a_{ij} = \begin{cases} a_i & \text{if }j = i\leq m \\ 0 & \text{if } j = i > m \\ 0 & \text{if }j \neq i \\ \end{cases}$. In other words, the only nonzero entries are on the diagonal, and satisfy this criterion. 

\thm

If $A$ is a $p \times n$ integer matrix, there exist invertible integer matrices $P, Q$, such that $PAQ$ is in Smith Normal Form.

\proof

We will use a modification of Gaussian elimination. Consider the matrix
\[
\begin{pmatrix} 
a_{11} & a_{12} & \cdots & a_{1n} \\
a_{21} & a_{22} & \cdots & a_{2n} \\
\vdots & \vdots & \vdots & \vdots \\
\end{pmatrix}
\]
Normally, we would take the 2nd row minus $\frac{a_{11}}{a_{21}}$ times the first. We aren't in a field, we are in $\Z$, so we can't do that. 

Instead, consider the $2\times2$ integer matrix $\begin{pmatrix} a & b \\ c & d \\ \end{pmatrix}$. Let $e = \gcd(a, c)$. We know $e = ax + cy$, for some $x, y \in \Z$. Now $a = e\alpha, c = e\beta$, then $1 = \alpha x + \beta y$. Then we have the invertible matrix
\[
\begin{pmatrix} x & y \\ - \beta & \alpha \\ \end{pmatrix} 
\]
Multiplying this matrix by $\begin{pmatrix} a & b \\ c & d \\ \end{pmatrix}$ yields a matrix with a $0$ in the bottom left. So, on the first column, we acted by 
\[
\begin{pmatrix} a \\ c \end{pmatrix} \mapsto \begin{pmatrix} e \\ 0 \end{pmatrix} 
\]
We combine this with column operations, which will be the same as the above row operations but transposed, and we can multiply $\begin{pmatrix} a & c \\ b & d \\ \end{pmatrix}$ by some matrix to yield a $2\times2$ matrix with a 0 in the upper right. 

We can upscale this algorithm, and use row and column operations, and we can do the thing we said we can do. In particular, at the end $a_{11}$ will be the $\gcd$ all matrix entries in the first row and column, etc. 

\qed

\cor

Any finitely generated ableian group is isomorphic to 
\[
\Z/a_1\Z \oplus \cdots \oplus \Z/a_m\Z \oplus \Z \underbrace{\oplus \cdots \oplus}_{t}\Z
\]
\qed

\section*{Lecture 10}

See handout on Gauchospace about Smith Normal Form. Sorry am bad

Apparently you can use Smith Normal Form to show that there are only 3 abelian groups of order 8...?

\subsection*{\underline{Groups of order $pq$, $p \neq q$ prime}}

Let $G$ have order $pq$. We know it has to contain a $p$-Sylow subgroup and a $q$-Sylow subgroup, both cyclic. 

\lem

If $(G:H)$ is the smallest prime dividing $|G|$, then $H\lhd G$. 

\proof

If $H = \Z/q\Z \subset G$, then $G/H \cong \Z/p\Z$. 

Consider $\{x \in G \mid xH$ has order $p\}$. What is $o(x)?$ Either $p$ or $pq$. If there is an $x$ with $o(x) = pq$, then $G \cong \Z/pq\Q$. Then we have an $x$ with $o(x) = p$, and $y$ with $o(y) = q$. Then $x^{-1}yx \in \langle y \rangle$. Then $x^{-1}yx = y^k$ for some $k$. So $x^{-p}yx^p = y^{kp}$. But $x^p$ is the identity. 

We need $kp \equiv 1\mod q$. 

$\Aut(\Z/q\Z) \cong (\Z/q\Z)^\times$ is cyclic of order $\phi(q) = q - 1$. So $q \mid (kp - 1 )$, so $kp - 1 \equiv 0 \mod q$, so $kp \equiv 1 \mod p$

Done? Maybe?

\qed

The point is $\Z/(pq)\Z \cong \Z/p\Z \oplus\Z/p\Z$. 

Let $G = \Z/p\Z \oplus\Z/q\Z$. We're gonna do ``the Smith Normal Form" thing, I guess. This group has generators $e_1 =(1, 0), e_2 = (0, 1)$. We have $pe_1 = qe_2 = e$, and $e_1e_2 = e_2e_1$. We want to do Smith Normal Form to the matrix $\begin{pmatrix} p & 0 \\ 0 & q \\ \end{pmatrix}$. We get $\begin{pmatrix} 1 & 0 \\ 0 & pq \\ \end{pmatrix}$. So there's only one generator...?

So we know what happens if $|G| = p$, $|G| = 4$, $|G| = pq$ with $p < q$. 

What if $|G| = 8,9,12,16,$, etc.? 

\lem

If $G$ is a $p$-group, then $G \supset\Z/p\Z$. 

\qed

So if $|G| = 8$, then $G$ contains a subgroup isomorphic to $\Z/2\Z$. By the lemma we proved earlier, this subgroup has to be normal. 









\end{document}